---
title: "Phen Flowering Time"
author: "Qingyuan Liu"
date: "2021/3/14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project aims to compare the flower duration of populations of Echinacea by sites in Western Minnesota.

```{r pacakages, message=FALSE, warning=FALSE}
library(here)
library(dplyr)
library(tidyr)
```

The raw data consists of two datasets: headIDtoAKA20190612.csv (explore) 
and remPhenologyThru2019_20200909.csv (explore2)

```{r}
explore <- read.csv(here('00_raw_data/headIdToAKA20190612.csv'))
explore2 <- read.csv(here('00_raw_data/remPhenologyThru2019_20200909.csv'))
head(explore)
head(explore2)
```

These two files contain the information of Echinacea populations, including their flower time. but We need to merge the two files into 1 file for further exploration.

```{r tidy, echo=TRUE, message=FALSE, warning=FALSE, paged.print=TRUE}
alldata <- merge(explore, explore2, by = "headID")
summary(alldata)
tidy <- (select(alldata, headID, site, year.y, startDtEarly,startDtLate, endDtEarly, endDtLate))
summary(tidy)
```

Since the flower duration need to be calculated from the start and end date, the date columns need to be separated into year, monnth and date.

```{r}
Duration <- tidy %>% 
  mutate(duration = as.Date(endDtEarly) - as.Date(startDtEarly))
summary(Duration)
```

Now dataset Duration is ready to be analyzed since it contains all the information needed, including the duration of flower time. 
But there are some columns containing "N/A", and some unavailable date that makes the result invalid.
We can clear all such columns first.
```{r}
nrow(Duration)
Dur_tidy <- Duration%>% 
  filter(duration > -1)%>%
  filter(duration < 50)%>%
  filter(site != "Other")%>%
  drop_na()
nrow(Dur_tidy)
```

All the sites contained in the raw data is shown below:

```{r}
unique(Dur_tidy$site)
```

Summarize the average flower duration according to sites:

```{r paged.print=TRUE}
Sum <- Dur_tidy%>%
  group_by(site)%>%
  summarise(avg_dur = mean(duration))
head(Sum)
```

Now we can start performing an One-way ANOVA test to see if the site has any impact on the flower time.
(Or if the flower time diverges in different sites.)

```{r}

```