---
title: "Markdown for phen project"
author: "Qingyuan Liu"
date: "2021/3/14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This project aims to compare the flower duration of populations of Echinacea by sites in Western Minnesota.

```{r pacakages}
install.packages("dplyr")
library(dplyr)
install.packages("tidyr")
library(tidyr)
```

The raw data consists of two datasets: headIDtoAKA20190612.csv (explore) 
and remPhenologyThru2019_20200909.csv (explore2)

```{r}
head(explore)
head(explore2)
```

These two files contain the information of Echinacea populations, including their flower time. but We need to merge the two files into 1 file for further exploration.

```{r tidy}
alldata <- merge(explore, explore2, by = "headID")
select(alldata, headID, site, year.y, startDtEarly,startDtLate, endDtEarly, endDtLate)
tidy <- (select(alldata, headID, site, year.y, startDtEarly,startDtLate, endDtEarly, endDtLate))
summary(tidy)
```

Since the flower duration need to be calculated from the start and end date, the date columns need to be separated into year, monnth and date.

```{r}
separated <- tidy %>%
  separate(startDtLate, into = c("yearz", "month", "day"), sep = "-") %>%
  separate(endDtEarly, into = c("yearz2", "month2", "day2"), sep = "-")
head(separated)
```
